# Stata {#sec-stata}

While working on the L2C main outcomes paper, I wanted to double check some of my results in Stata. It turns out you can run Stata code directly from within R; albeit with some limitations. This note needs some work, but the basics are here.

```{r}
#| echo: false

# Silently load the mtcars data
data("mtcars")

# For testing: Export the mtcars data to desktop for testing in Stata.
# haven::write_dta(mtcars, "/Users/bradcannell/Desktop/mtcars.dta")
```

The first thing we need to do is load the `RStata` package.

```{r}
library(dplyr, warn.conflicts = FALSE)
library(RStata) # For Stata
```

Then, we need to tell R where to find Stata on our local computer and which version of Stata we are using. 

```{r}
# For Stata
options("RStata.StataPath" = "/Applications/Stata/StataMP.app/Contents/MacOS/stata-mp")
options("RStata.StataVersion" = 18)
```

Then, we can run Stata directly in R code chunks in RStudio. For example, let's calculate the frequency and percentage of observations of each cylinder type in the `mtcars` data set.

1. Pass a Stata command (as a character string) -- or a do file -- to the `src` argument of the `stata()` function.
2. Pass the name of a data frame that currently exists in the R global environment to the `data.in` argument of the `stata()` function.

```{r}
stata(
  src = "tabulate cyl",
  data.in = mtcars
)
```

This works, but we can't use these estimates directly. However, we can save the estimation results and use them in further operations. See this [SO post](https://stackoverflow.com/questions/57881623/how-to-capture-rstata-package-output-for-subsequent-processing-in-r) for more discussion. 

```{r}
# Create a temp file to save the results to.
# I don't want to create a dta file on my desktop every time I build this book.
tmpdir <- tempfile()
dir.create(tmpdir, showWarnings = FALSE, recursive = TRUE)
tempfile <- file.path(tmpdir, "stata_out.dta")
```

```{r}
stata_tabulate_results <- stata(
  paste0( # Using past to inject the temp file. Normally this isn't necessary
    '
    tabulate cyl, matcell(x)
    clear
    svmat x
    ',
    'save ', tempfile
  ),
  data.in = mtcars,
  data.out = TRUE
)
```

Without the tempfile code, it would look like this.

```{r}
#| eval: false

stata_tabulate_results <- stata(
  '
  tabulate cyl, matcell(x)
  clear
  svmat x
  save "/Users/bradcannell/Desktop/t.dta"
  ',
  data.in = mtcars,
  data.out = TRUE
)
```

I still have a lot to learn about saving the results from Stata commands. In the future, we may want to add notes from "/Users/bradcannell/Library/CloudStorage/Dropbox/Stata/Notes/Example Do Files/Matrix and Saved Results.do"

```{r}
#| label: stata-clean-up
#| echo: false

rm(list = ls())
```
